---
title: "Bears"
author: "MITCH"
date: "12/14/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```





## Question


Ranger Bob Berra needs help relocating bears out of Jellystone. He wants to move bears with higher weights but doesn't want to go and weight every bear. He asked to find a better way to determine the bears weight.

With this presentation I will demonstrate what I know and learned from the class


##

```{r bears}
library(DT)
library(effects)
library(car)
library(readr)
library(ggplot2)
library(tidyverse)
bears <- read_csv("/home/students/holmanma/Des_Models/Models/bears.csv")

bears %>%
  group_by(sex)%>%
  summarise(mean (chest.g), mean(neck.g))


```


##


```{r, echo=FALSE}
names(bears)
modfo <- lm(weight ~ . , data = bears)
summary(modfo)
modfo
```


##


```{r}
ggplot(data = bears, aes(x = weight, y =  neck.g )) + 
  geom_point() + 
  theme_bw() + 
  stat_smooth(method = "lm", formula = y ~ poly(sqrt(x), 1))
```

##

```{r}
ggplot(data = bears, aes(x = weight, y = chest.g  )) + 
  geom_point() + 
  theme_bw() + 
  stat_smooth(method = "lm", formula = y ~ poly(sqrt(x), 4)) 
```


##


```{r}
ggplot(data = bears, aes(x = weight, y =  chest.g )) + 
  geom_point() + 
  theme_bw() + 
  stat_smooth(method = "lm", formula = y ~ poly(x, 1))
```



## What I have learned

I can make a website using R studio and git hub.

First

open and create a text file save it _site.yml

Second 

name the site from your project

 name: "Mitchell-Holman.github.io"
output_dir: "."

##


Third 

create navbar
navbar:

  title: "Mitchell Holman"
  left:
  - text: "Bike Project"
    href: Projects.html
  - text: "Hobbies"
    href: hobs.html
  - text: "Time Series"
    href: Time.html
  right:
  - icon: fa-github fa-lg
    href: https://github.com/Mitchell-Holman/Mitchell-Holman.github.io
output:
  html_document:
    theme: yeti
    
now build your site

##
