---
title: "lab7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.	Generate an approximate AR(1) process y_t with 0 mean and constant variance by first randomly generating the white noise ε_t for t=1,2,3, …,100 then using a truncated version of the MA representation of an AR(1) process for some value of ϕ, where -1< ϕ<1.  Specify the ϕ value you used.  Attach and describe the time series plot.

```{r}
set.seed(42)
t <- 1:100
et <- rnorm(t, 0, 1)
AR1 <- arima(et,c(1,0,0))
```

```{r}
phi <- 0.4
y <- c()
k <- 1:length(et)
for (i in k){
  y[i] <- et[i] + et[i-1]*phi + et[i-2]*phi^2 + et[i-3]*phi^3 + et[i-4]*phi^4 + et[i-5]*phi^5
    }
yt <- y[6:100]
plot.ts(yt)
```


2.	Attach and describe the sample ACF and the sample partial ACF of the time series.  
```{r}
ayt <- acf(yt,95)
pacf(yt,95)
```
Because the data was generated using an AR(1) process it makes sense that the ACF shows more of a pattern than the partial ACF. Each value was calculated using the previous 5 values and the partial ACF ignores the effect of the intermediate values between the lags.

3.	Do the plots in #2 appear consistent with an AR process?  Explain.
Yes they do, the correlation increases and decreases about every 5 values because each value was calcuated using the previous 5 values.



4.	Using the ϕ value in #1, calculate the theoretical lag k autocorrelation of y_t  for k =1,2,3 if  y_t=〖ϕy〗_(t-1)+ε_t. 

```{r}
rho1 <- phi^1
print(rho1)
rho2 <- phi^2
print(rho2)
rho3 <- phi^3
print(rho3)
```



5.	Compare the values in #4 with the corresponding values of the sample autocorrelation.  

```{r}
ayt[1:3]

```

6.	Investigate and discuss the effects of changing the value of the parameter ϕ of the AR(1) process on the theoretical Var(y_t) and Cov(y_t,y_(t+1) ).  

```{r}

```

7.	Are these effects in #6 reflected on the sample ACF?  Discuss.

```{r}

```


8.	Fit an AR(1) model to the time series y_t  generated in #1.  What are the estimates of the model parameters? Is this model appropriate for the data?  Explain.

```{r}

```


9.	Obtain and discuss the sample ACF and sample partial ACF of the residuals after fitting the model.  

```{r}

```


10.	Fit an MA(1) model to the time series y_t  generated in #1.  What are the estimates of the model parameters?   Is this an appropriate model for the data? Explain.

```{r}

```



Upload the completed .html file at the Activity 7 Submission folder in AsULearn by midnight on Tuesday, Nov. 7, 2017.
